---
title: "Homework 2"
output: 
  html_document:
    self_contained: yes
---
---
title: "Homework 2"
output: 
  html_document:
    self_contained: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T,eval=T,message=F,warning=F,fig.align='center')
library(tidyverse)
source("../scripts/ggprob.R")
```


## Problem 1 <small>(2pts each, 8pts total)</small>

Here are a few probability exercises to get you warmed up.

a. Consider $X\sim \text{Normal}(5, 2^2)$. Find $P(4 < X < 7)$
```{r}
m = 5
sd = 2

ggplot() + geom_norm_density(mu=m, sigma=sd, color = "black") + 
  geom_norm_fill(m, sd, a=4, b=7, fill = "red", alpha = 0.5)

pnorm(7, m, sd) - pnorm(4, m, sd)

```
**Answer**: $P(4 < X < 7) = 0.5328072$ 

b. Suppose you have an uneven 6-sided die where the even numbers are equally likely to occur and the odd numbers are also equally likely. The expected value of the die is $3.3333$ (i.e. $\frac{10}{3})$. What is the probability of getting a 6?
   (Hint: Let $p$ represent the probability of getting an even number, and represent the probability of the other numbers algebraically. Then derive an expression for the expected value, let it equal $\frac{10}{3}$, and solve. You can go back and check your work by plugging the result back into your original expressions and verify it works with R.)

**Answer**: The probability of getting a $6$ is $\frac{1}{9}$.

**My work**: Let $X$ be a random variable that can take on values $\{1, 2, 3, 4, 5, 6\}$ via the random process of a die roll. Let $p$ represent the probability of getting an even number. Because we are given that the even numbers are equally likely to occur, and there are only 3 options for even numbers, we know that $P(2)=P(4)=P(6) = \frac{p}{3}$. Let $o$ represent the probability of getting an odd number. By the same reasoning, we know that $P(1)=P(3)=P(5)=\frac{o}{3}$. We know that all probabilities must add up to one. Thus we know $p + o = 1$. Now we must use the general formula for the expected value: $\sum_{i=1}^{6}x_i *P(X=x_i)$ where $x_1 = 1$, $x_2 = 2$, etc. Expanding this summation and combining like terms gives us $9o + 12p = 10$. Now we have two linear equations involving $o$ and $p$. When we solve, we get that $p = \frac{1}{3}$. We want to find $P(6)$. So we get $P(6) = \frac{p}{3} = \frac{\frac{1}{3}}{3} = \frac{1}{9}$. 


c. Give **two examples** of pairs of events that are mutually exclusive and explain why for each.

**Answer**: 

1. Say you are flipping a coin once. Let A be the event of observing a Head and B be the event of observing a Tail. A and B are mutually exclusive because their intersection is the empty set. This is because with one coin flip it is not possible to observe both a Head and a Tail, only one or the other can occur.  
2. Suppose you roll a six sided fair die once. Let C be the event that you observe an even number and D be the event that you observe an odd number. C and D are mutually exclusive events because a number can only be even or odd, not both. Thus, C intersection D is the empty set.

d. Give **two examples** of pairs of events that are independent and explain why for each.

**Answer**: 

1. Say you roll a fair six sided dice twice. Let A be the event that you observe an even number on the first roll and B be the event that you observe an odd number on the second roll. $P(A) = 0.5$ and $P(B) = 0.5$. $A \cap B$ is the even that you get an even number on the first roll **and** an odd number on the second roll. The outcome space of rolling the dice twice is $\{EE, EO, OE, OO\}$. Thus $P(A \cap B) = \frac{1}{4}$. Because $P(A \cap B) = P(A) * P(B)$ ($0.25 = 0.5 * 0.5$), A and B are independent events. More generally this means that A and B are events that occur without being impacted by each other. 
2. Say there is a hat that has 20 scraps of paper in it, each with an integer between 1 and 20 written on it. Suppose you draw one scrap from the hat. Let C be the event that the number on the paper is divisible by 5. That is, $C = \{5, 10,15,20\}$. Let D be the event that the number is odd. That is, $D = \{1,3,5,7,9,11,13,15,17,19\}$. Since there are 20 overall possible outcomes, $P(C) = \frac{1}{5}$ and $P(D) = \frac{1}{2}$. Now we must investigate the event $C \cap D$, which is the event that the number is both divisible by 5 **and** odd. Thus, $C \cap D = \{5, 15\}$ and $P(C \cap D) = \frac{1}{10}$. Notice that $P(C) * P(D) = \frac{1}{10} = P(C \cap D)$, thus C and D are independent events. 

## Problem 2 <small>(2pts each, 14pts total)</small>

For each of the following scenarios, decide what random variable may be appropriate to use as a model, and explain your reasoning as well as any assumptions you make.

a. The number of patients in an experimental drug trial that respond to treatment. 

**Answer**: I think a Binomial RV would be appropriate to use to model this situation. As for the parameters, n would be the number of patients that the trial is administers to, and p would be the probability that a trial is a success (aka that a patient responds to treatment). To satisfy the BINS assumptions we must assume the following: that each trial corresponds to a patient that hasn't before done the treatment, the outcomes are binary (B: either a patient responds to treatment or they don't), each trial is independent of the others (I: one patient's response doesn't impact another patient's), there are a fixed number of trials (N), and each patient has the same probability of responding to treatment (S).  

b. The number of red cars you observe driving through an intersection between 10am and 11am. 

**Answer**: I think a Poisson RV would be appropriate to model this situation. For this model to work we must make the assumption that there exists a positive, real number called lambda that represents average rate of red cars that one observes driving through a specific intersection between 10 and 11am.  

c. What the second hand reads on the clock when you wake up from a long nap (approximately). 

**Answer**: I think a Discrete Uniform RV would be good to model this situation. The second hand reads the minute past the hour on a clock. Thus, this RV models approximately what minute after the hour it is after you wake up from a long nap. There are exactly 60 minutes in an hour, and a clock's second hand doesn't accurately represent any unit of time smaller than a minute. Thus, our possible outcomes are the integers between 0 and 59, inclusive. So if using a discrete uniform RV, the parameters a and b would be 0 and 59 respectively. We must make the assumption that if you wake up and the second hand is not exactly on a minute marker, then you must round to the nearest whole minute. This model should work because each minute has the same constant probability of occurring, that is 1/60.  


d. How many people you need to swipe right on Tinder before you get a match. 

**Answer**: I think that a Geometric RV would be good to model this situation. Here the RV would represent the number of failures (swipe rights) before you get a success (a match). In order for the Geometric model to work, we must make the assumption that each trial (person that you view on Tinder), is independent of previous trials. We must also assume that each trial has the same probability of a successful match. 

e. The length of time between mosquito bites a camper experiences while on a hike. 

**Answer**: I think that an Exponential RV would be good to model this situation. The RV would model the waiting time that occurs between each mosquito bite (event) during a the period of time that a camper is on a hike. In order for this model to work we must assume that there is a rate lambda that represents the expected wait time between bites and that an occurrence of a single bite is not dependent on previous bites.  

f. Whether the Eagles win the Superbowl this year. 

**Answer**: I think that a Bernoulli RV would be a good way to model this situation. This is because Bernoulli RV's deal with 1 trial where there are only two outcomes. Here our single trial would be the Superbowl game and the outcomes are either that the Eagles win or they don't. We must assume that there exists some p that represents the probability that the Eagles do win the Superbowl this year. 

g. The GPA of a randomly selected UW-Madison graduate.

**Answer**: I think that a Continuous Uniform RV would work well for modeling this situation. This is because there is a closed continuous interval possible of GPAs that a random UW-Madison graduate could have: [0, 4]. For this model to be accurate we must assume that the density is constant for each outcome. Using the interval [0,4], the constant density would be $\frac{1}{4}$. I could also see the argument for using a Normal RV to model this situation, however the outcome space of a Normal RV is $(-\infty, \infty)$. GPA can only between 0 and 4, so a Normal RV wouldn't technically work.  


## Problem 3 <small>(2pts each, 4pts total)</small>

a. Consider the CDF given below for random variable $Y$ 
```{r}
plot(x=0:6, y=c(0,.3,.5, .6, .75,.8,1), type="s", ylab="F(y)", xlab="y", main="cdf of Y")
abline(h=seq(.1,.9,.1), lty=2, col="gray")
```
Express the pmf of $Y$ in the following R code (in other words, fill in the missing probabilities)

```{r}
y<- 1:6
py <- c(0.3,0.2,0.1,0.15,0.05, 0.2) ## edit this line
barplot(height=py, names=y, main="pmf of Y")
```



b. Calculate $E[Y]$ for the variable $Y$ in part a.

```{r}
expected_value = sum(y * py)
expected_value
```
**Answer**: $E[Y] = 3.05$.


## Problem 4 <small>(2pts each, 4pts total)</small>

Below you will find the plot of a probability density function for a random variable $X$.

```{r, echo=FALSE}
plot(x=c(0,1,2,3), y=c(0,.5,.5,0), type="l", ylab="density", xlab="x", main="pdf of X")
```

Using just geometry answer a and b:

a. What is $Pr[X < 1]$?

**Answer**: $Pr[X < 1] = 0.25$ 

b. What is $Pr[1.5 < X < 2.5]$
```{r}
7/16
```


**Answer**: $Pr[1.5 < X < 2.5] = 0.4375$



**Answer**: $Pr[1.5 < X < 2.5] = 0.4375$


